### 2015-04-20 Paper2015.tex
Incorporated second page of comments from Alexander.

% Paper 2015
\documentclass[peerreviewca,11pt,a4paper]{IEEEtran}

\usepackage{ifpdf}
\usepackage{cite}

\ifCLASSINFOpdf
    \usepackage[pdftex]{graphicx}
    \graphicspath{{../pdf/}{../jpeg/}}
    \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
    \usepackage[dvips]{graphicx}
    \graphicspath{{../eps/}}
    \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[cmex10]{amsmath}

\usepackage{array}
\usepackage{fixltx2e}

\ifCLASSOPTIONcaptionsoff
  \usepackage[nomarkers]{endfloat}
 \let\MYoriglatexcaption\caption
\renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
\fi

\usepackage{url}

\usepackage{siunitx}
\DeclareSIUnit\photon{\ensuremath{\gamma}}
\DeclareSIUnit\proton{p}
\DeclareSIUnit\neutron{n}
\DeclareSIUnit{\seivert}{Sv}

\usepackage{booktabs}

\usepackage{graphicx}
\usepackage{epstopdf}

\usepackage{caption}
\usepackage{subcaption}

\usepackage{todo}
%\renewcommand\todo[1]{\relax}
%\renewcommand\todos{\relax}

\usepackage{bpchem}
\def\U238{\BPChem{\^{238}U}}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{
    Detailed Geant4 simulations of the ANITA and ANITA-CUP neutron facilities
}

\author{
    \IEEEauthorblockN{
        Q. Hong\IEEEauthorrefmark{1},
        S. P. Platt\IEEEauthorrefmark{1},
        A. V. Prokofiev\IEEEauthorrefmark{2} and
        E. Passoth\IEEEauthorrefmark{3}
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{1}
        University of Central Lancashire, England
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{2}
        Division of Applied Nuclear Physics and The Svedberg Laboratory, University of Uppsala, Sweden
    }
    \IEEEauthorblockA{
        \IEEEauthorrefmark{3}
        The Svedberg Laboratory, University of Uppsala
    }
}

\maketitle

\begin{abstract}
    Monte Carlo simulation of spallation neutron source of ANITA at TSL was used to analyze radiation fields at the Close User Position (CUP) and Standard User Position(SUP) for single-event effect testing. ...
    \todo{Revise abstract}
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}

The ANITA facility~\cite{Prokofiev2009} at The Svedberg Laboratory (TSL) in Uppsala, Sweden, has been widely used for studies and accelerated testing for neutron single-event effects (SEE).
A high-flux upgrade of the facility, ANITA-CUP, has recently been introduced and reported by Prokofiev et al.~\cite{Prokofiev2014}.

Platt et al. have reported preliminary Monte Carlo simulations of neutron and gamma production at the standard ANITA facility~\cite{Platt2013}.
That work was motivated by a need to understand possible sensitivity to gamma rays in detectors used for beam monitoring during SEE tests.
Simulations of a bare spallation target showed qualitative agreement with independent calculations and measurements of the fast neutron field at the standard ANITA facility~\cite{Prokofiev2009}.
The present work was motivated partly by a desire to investigate whether improvements in the fidelity of the geometry used in the simulations would lead to closer quantitative agreement with independent data, and partly by a need to extend the work to encompass the new ANITA-CUP facility~\cite{Prokofiev2014}.

Accordingly, in this paper we present a comprehensive set of analyses of neutron and gamma fields at ANITA and ANITA-CUP, with comparison to independent models and validation against measurements.
We present a subset of those results in this extended abstract.

\section{Modelling}
Simulations were conducted using Geant4 version 10.0~\cite{Agostinelli2003,Allison2006} with the binary intranuclear cascade model (see also~\cite{Platt2013}).
A detailed model of the ANITA geometry~\cite{Prokofiev2009,Prokofiev2014} was implemented in Geant4, as illustrated in Fig.~\ref{fig:ANITAoverview}.
%; a summary only is given here.
%The spallation target is a tungsten cylinder of diameter \SI{5}{\cm} and length \SI{2.4}{\cm}~\cite{ANITAdrawing}.
%The target is cooled by water and surrounded by a stainless-steel cooling jacket.
%The target assembly with its cooling jacket appears as a small blue rectangle in Fig.~\ref{fig:ANITAoverview}.
%Lead shielding blocks in the target region are also modelled, and shown in red Fig.~\ref{fig:ANITAoverview}.
%A large electromagnet surrounding the target is modelled and shown in Fig.~\ref{fig:ANITAoverview} in grey (iron) and yellow (copper).
%An iron collimator with variable apertures is modelled  and shown downstream of the target (grey).
Simulated protons were incident axially at an energy of \SI{180}{\MeV}.
Resulting neutron and gamma fields were evaluated at several detector locations.
In this paper results are presented for three of these: the Standard User Position (SUP), \SI{2.5}{\m} downstream of the target, the Close User Position (CUP), \SI{0.75}{\m} downstream of the target, and the CUP-TOF position, \SI{0.84}{m} downstream of the target.
%N.B. This is downstream of the downstream edge of the target in Qian's model.
The SUP position is downstream of a frontal wall with a collimator with variable aperture; the CUP and CUP-TOF positions are upstream of the same collimator.
The SUP and CUP positions are of interest as they are positions used for SEE tests; the CUP-TOF position is the location of thin-film breakdown counter (TFBC) detectors used for beam monitoring and characterisation including neutron time of flight (TOF) measurements.
TFBC data are used to validate simulation results.

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{overview.png}
	\caption{
        Simulated ANITA facility overview seen from above (yellow: bending magnet; red: shielding components; blue: neutron production target assembly; grey: collimator; blue: virtual detector system)
    }
	\label{fig:ANITAoverview}
\end{figure}

\section{Results}

\subsection{Neutrons}
Fig.~\ref{fig:SUPDensity} shows neutron distribution at the SUP, for neutrons above \SI{10}{\MeV}.
The collimator ensures that the beam is circular. Within the beam umbra the calculated neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{7.1e5}{\neutron\per\cm\squared\per\second}, compared to \SI{5.7e5}{\neutron\per\cm\squared\per\second} from earlier calculations~\cite{Platt2013} and \SI{9.3e5}{\neutron\per\cm\squared\per\second} from measurements.
The measurements therefore exceed the present calculations by about 30\%.
\todo{Check SUP flux.}

\begin{figure}[t]
    \begin{minipage}{\columnwidth}
        \includegraphics[width=0.9\columnwidth]{SUP10ColSpatialDistribution10MeVRADECS.png}
        \subcaption{
            Standard User Position.
            Concentric white circles indicate nominal umbra and penumbra limits
        }
        \label{fig:SUPDensity}
        \todo{
            Fig.~\ref{fig:SUPDensity} replace with 1D profile for neutrons above \SI{10}{\MeV}.
             Show position of collimator.
        }
    \end{minipage}
    \begin{minipage}{\columnwidth}
        \includegraphics[width=0.9\columnwidth]{CUP10ColSpatialDistribution10MeV.png}
        \subcaption{Close User Position}
        \label{fig:CUPDensity}
    \end{minipage}
    \caption{
        Spatial distribution of neutrons above \SI{10}{\MeV} at SUP and CUP, with \SI{10.2}{\cm} diameter collimator
    }
\end{figure}

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.9\columnwidth]{CUPTOF10beamproRADECS.png}
    \caption{
        Vertical profile of \U238\ fission events at the CUP, with \SI{10.2}{\cm} diameter collimator. }
    \label{fig:CUPProfile}
    \todo{
        Fig.~\ref{fig:CUPProfile}.
        Revise legend.
        Plot lines +symbols.
        Add vertical lines?
    }
\end{figure}

Fig.~\ref{fig:CUPDensity} shows the spatial distribution of neutrons above \SI{10}{\MeV} at the CUP.
As CUP is upstream of the collimator the distribution is not not circularly symmetric.
In the central region the calculated neutron fluence rate above \SI{10}{\MeV} at a proton current of \SI{200}{\nA} is \SI{7.85e6}{\neutron\per\cm\squared\per\second}, compared to \SI{1.17e7}{\neutron\per\cm\squared\per\second} from measurements~\cite{Prokofiev2014}.
The measurements therefore exceeds the present calculations about 50\%.
\todo{Check current consistency -- make \SI{215}{\nA}, check Elke's report.}

TFBCs equipped with \U238\ targets were used to measure beam profile and time of flight spectra at the CUP-TOF position.
Simulated neutron counts were folded in energy with the \U238\ fission cross-section~\cite{Carlson2009} and compared with measurements.
Fig.~\ref{fig:CUPProfile} shows the resulting vertical distribution of the neutron-induced fission rate in \U238.
The results show close agreement between measurement and simulation, although the sharp cut-off beyond \SI{11}{\cm} is not fully reproduced by the simulation.%
\todo{
    Improve presentation of results for CUP profile.
    Make calculations at the same positions as those used for the measurements.
}

In Fig.~\ref{fig:TOFSpectra} we compare simulated and measured TOF spectra at the CUP-TOF position.
Neutron TOF data from the simulation were folded in energy with the \U238\ fission cross-section, convolved with a \SI{5}{\ns} rectangle function approximating the primary proton micropulse shape, and overlapped at \SI{45}{\ns}, representing the timing ambiguity due to the micropulse period.
The results show very close agreement in the position and width of the TOF peak.
The continuum beyond about \SI{30}{\ns}, corresponding to neutrons below about \SI{40}{\MeV}, is somewhat higher in the calculation than in the measurement.
This might indicate a relative overestimate of lower energy neutrons in the calculations -- or an corresponding underestimate of higher-energy neutrons.
Prokofiev et al.~\cite{Prokofiev2014} note that the TOF spectrum at the CUP comprises three parts, a component directly from the target, one scattered forward from target surroundings, and one scattered backwards from the front face of the collimator block.
The present simulations confirm the Prokofiev et al. semi-empirical multicomponent model. Details of this confirmation will be presented in the full paper but are omitted here for brevity.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\columnwidth]{CUPTOFtofspectraRADECS.png}
    \caption{
        Time of flight spectra for \U238 fission events at the CUP-TOF position, with \SI{3}{\cm} diameter collimator.
        }
    \label{fig:TOFSpectra}
    \todo{
        Figure:~\ref{fig:TOFSpectra}: Edit legend, edit y label, make simulated data a curve.
    }
\end{figure}

\begin{figure}[t]
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.8\columnwidth]{SUPNormalisedRADECS.eps}
        \subcaption{
            SUP.
            The comparison is against the fit of Prokofiev et al.~\cite{Prokofiev2009} and preliminary Geant4 modelling by Platt et al.~\cite{Platt2013}
        }
        \label{fig:LethargyplotSUP}
    \end{minipage}
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.8\columnwidth]{CUPcomparedProkofiev2014NormalisedRADECS.eps}
        \subcaption{
            CUP.
            The comparison is against the fit of Prokofiev et al.~\cite{Prokofiev2014}
        }
        \label{fig:LethargyplotCUP}
    \end{minipage}
	\caption{
        Calculated neutron spectra, normalised to fluence above \SI{10}{\MeV}, with
        \SI{10.2}{\cm} diameter collimator
    }
	\label{fig:Lethargyplots}
\end{figure}

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.8\columnwidth]{DiffYieldComparedSUPCUP10.eps}
    \caption{
        Differential neutron spectra at CUP and SUP (\SI{10.2}{\cm} diameter collimator), comparing results of this work to analytical models~\cite{Prokofiev2009,Prokofiev2014}
    }
    \label{fig:DifferentialSpectra}
\end{figure}

Fig.~\ref{fig:Lethargyplots} compares calculated spectra at SUP and CUP with their respective analytical models.
The SUP data are also compared with results from preliminary Geant4 simulations~\cite{Platt2013}.
These curves are normalised to neutron fluence above \SI{10}{\MeV} an presented as lethargy plots to emphasis the shape of the spectra rather than their integral fluence rate.
The increased integral fluence rate at SUP from the present calculations, compared to the preliminary results of \cite{Platt2013} is therefore not visible in Fig.~\ref{fig:LethargyplotSUP}.
Fig.~\ref{fig:LethargyplotSUP} does show an increased contribution below about \SI{1}{\MeV}, consistent with interactions in materials other than the target, which were absent from the preliminary simulations.
Fig.~\ref{fig:LethargyplotCUP} shows extremely close agreement with the analytical model from~\cite{Prokofiev2014} (the analytical model smooths out the structure in the evaporation peak around \SI{1}{\MeV}).
When compared to the analytical models the calculated spectra from the current work lead to slightly lower neutron yield and slightly higher peak energies in the higher energy peak near \SI{100}{\MeV}.
It is not surprising that this is not affected by the improved model geometry as this peak arises from intranuclear cascades in the target.
We continue to investigate this discrepancy and will discuss it further in our final paper.

Fig.~\ref{fig:DifferentialSpectra} summarises neutron spectra results by comparing differential flux as calculated in this work with the Prokofiev et al. models~\cite{Prokofiev2009,Prokofiev2014}.
The SUP calculations and analytical model agree very closely below \SI{20}{\MeV}; the analytical model exceeds the calculations somewhat above that energy.
At CUP the analytical model exceeds the Geant4 calculations by about 50\%, slightly more than that above \SI{20}{\MeV}.

\subsection{Photons}
Fig.~\ref{fig:GammaSpatialDistribution} shows the calculated spatial distribution of gamma photons at SUP and CUP, and shows a gamma radiation distribution that is qualitatively very similar to that for neutrons at both locations.
Fig.~\ref{fig:DifferentialGammaSpectra} shows the calculated differential gamma spectra within the central region of the spatial distribution.
There is a broad continuum up to an energy of about \SI{10}{\MeV}, with a peak in the fluence spectrum at about \SI{0.5}{\MeV} (where the annihilation peak is visible).
The gamma fluence rate is an order of magnitude greater at CUP that at SUP, consistent with the closer range to the target and its immediate surroundings.
The CUP spectrum also shows a backscatter peak around \SI{0.2}{\MeV}.

\begin{figure}[t]
    \begin{minipage}{\columnwidth}
        \includegraphics[width=0.9\columnwidth]{SUP10ColSpatialDistributionAllG.png}
        \subcaption{Standard User Position}
        \label{fig:GammaSpatialDistributionSUP}
    \end{minipage}
    \begin{minipage}{\columnwidth}
        \includegraphics[width=0.9\columnwidth]{CUP10ColSpatialDistributionAllG.png}
        \subcaption{Close User Position}
        \label{fig:GammaSpatialDistributionCUP}
    \end{minipage}
    \caption{
        Calculated gamma spatial distribution, with
        \SI{10.2}{\cm} diameter collimator
    }
    \label{fig:GammaSpatialDistribution}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\columnwidth]{gDYieldcomparedRADECS.eps}
    \caption{
        Calculated gamma spectra at CUP and SUP, with
        \SI{10.2}{\cm} diameter collimator
    }
    \label{fig:DifferentialGammaSpectra}
\end{figure}

\begin{figure}[t]
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{DoseVSenergySUP.eps}
        \subcaption{Standard User Position}
        \label{fig:GammaDoseEnergySUP}
    \end{minipage}
    \begin{minipage}{\columnwidth}
        \centering
        \includegraphics[width=0.9\columnwidth]{DoseVSenergyCUP.eps}
        \subcaption{Close User Position}
        \label{fig:GammaDoseEnergyCUP}
    \end{minipage}
    \caption{
        Calculated gamma dose spectra, with
        \SI{10.2}{\cm} diameter collimator
    }
    \label{fig:GammaDoseEnergy}
\end{figure}

Fig.~\ref{fig:GammaDoseEnergy} shows the calculated distribution of gamma dose, obtained by folding the gamma spectra (Fig~\ref{fig:GammaDoseEnergy}) with dose conversion data from \cite{Kwon1980}.
The results show structure with peaks around \SI{1.5}{\MeV}, \SI{3}{\MeV}, \SI{6}{\MeV} and \SI{9}{\MeV}.
Note that these are not coincidence peaks, as the simulation counts individual photons.
The calculated dose rate at \SI{200}{\nA} primary proton current is \SI{21}{\milli\seivert\per\hour} at SUP and \SI{370}{\milli\seivert\per\hour} at CUP.
The calculated value at SUP is slightly more than and therefore consistent with that predicted from a bare spallation target (\SI{17}{\milli\seivert\per\hour}) and is also consistent with an estimated upper limit reported in~\cite{Prokofiev2009} (\SI{40}{\milli\seivert\per\hour}).
The calculated value at CUP is consistent with that calculated at SUP (a ratio of 17:1 is somewhat greater than $1/R^2$, as would be expected), but rather more than the estimated upper limit given in \cite{Prokofiev2014} (\SI{170}{\milli\seivert\per\hour}).
We continue to evaluate our gamma calculations and will discuss them at greater length in the final paper.

\section{Discussion}
\todo{Discussion}

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem{Prokofiev2009}
Prokofiev, A. V. et al.,
\newblock ``Characterization of the ANITA neutron source for accelerated SEE testing at The Svedberg Laboratory'',
\newblock {\em 2009 IEEE Radiat. Effects Data Workshop Record}, pp.~166-173

\bibitem{Prokofiev2014}
Prokofiev, A. V. et al.,
\newblock ``CUP--A new high-flux irradiation position at the ANITA neutron facility at TSL'',		\newblock {\em IEEE Trans. Nucl. Sci.} vol.~61 no.~4, pp.~1929-1936, 2014

\bibitem{Platt2013}
Platt, S. P. et al.,
\newblock ``Neutron and gamma fields at neutron spallation sources for single-event-effects testing'',
\newblock {\em Proc. 14th European Conf. Radiat. Effects Compon. Syst.}, 2013

\bibitem{Agostinelli2003}
Agostinelli, S. et al.,
\newblock``Geant4--a simulation toolkit'',
\newblock {\em Nucl. Inst. Meth. Phys. Res. A} vol.~506, pp.~250--303, 2003

\bibitem{Allison2006}
Allison, J. et al.,
\newblock ``Geant4 developments and applications'',
\newblock {\em IEEE Trans. Nucl. Sci.} vol.~53 no.~1, pp.~270--278, 2006

\bibitem{Carlson2009}
Carlson, A. D. et al.,
\newblock ``International evaluation of neutron cross section standards'',
\newblock{\em Nucl. Data Sheets} vol.~110 no.~12, pp.~3215-3324, 2009

\bibitem{Kwon1980}
Kwon, S.-G. et al.,
\newblock ``Calculation of neutron and gamma-ray flux-to-dose-rate conversion factors'',
\newblock{\em J. Kor. Nucl. Soc.} vol.~12 no.~3, pp.~171-179, 1980

\end{thebibliography}

\cleardoublepage

\todos

\end{document}